TersectBrowser+ : A single browser for in-depth introgression analysis of re-sequenced genome datesets

### 2025-04-28 Article/Article.tex
Article Draft ready for Tom

\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}

@@

@@

\section{Features}
Tersect Browser+ is a web application that depends on resequenced data in VCF format and uses a reference genome in FASTA format. Utilizing a MEAN stack (MongoDB, Express, Angular, and Node) architecture, it creates a webpage displaying heatmaps and phylogenetic trees for introgressions within the loaded crop dataset. This provides an interface for users to identify differences across accessions within the same species dataset. It also includes a JBrowse interface for investigating specific variants, displaying selected genes directly alongside the generated heatmap. A search feature highlights high-impact variants on the webpage, helping to distinguish the selected reference genomes. Tersect Browser+ also offers barcode generation functionality for various cultivars and specific species in the dataset.

\begin{figure} [http]
  \centering
  \includegraphics[width=\textwidth]{path/to_image.png}
  \caption{Tersect Browser+ overview. The browser provides a single way to analyse variant data without leaving the page.[More description]}
  \label{fig:your_label}
\end{figure}

\subsubsection{Gene Brwoser}

We will include text here on the feature. I will look through the Technical Documentation to put this together.

\subsubsection{Gene Model}

We will include text here on the feature. I will look through the Technical Documentation to put this together.

We will include text here on the feature. I will look through the Technical Documentation to put this together.

We will include text here on the feature. I will look through the Technical Documentation to put this together.

@@

@@

A question for Tom - we see this included in many papers published in this journal. What should we include?

%\section{This is an example for first level head}\label{sec3}

% \subsection{This is an example for second level head - subsection head}\label{subsec2}

% \subsubsection{This is an example for third level head - subsubsection head}\label{subsubsec2}

% \paragraph{This is an example for fourth level head - paragraph head}

% \section{Conclusion}

% Some Conclusions here.

%%%%%%%%%%%%%%

\section{Code availability}

We will add in a link to the github reference in here and a link to the technical documentation.

>>>>

\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}%
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%
\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%
\usepackage{hyperref}%

@@

% EARLIER TEXT KEEP UNTIL SAFE TO DELETE

@@

\section{Features}

@@

Tersect Browser+ is a web application that depends on resequenced data in VCF format and uses a reference genome in FASTA format. Utilizing a MEAN stack (MongoDB, Express, Angular, and Node) architecture, it creates a webpage displaying heatmaps and phylogenetic trees for introgressions within the loaded crop dataset. This provides an interface for users to identify differences across accessions within the same species dataset. It also includes a JBrowse interface for investigating specific variants, displaying selected genes directly alongside the generated heatmap. A search feature highlights high-impact variants on the webpage, helping to distinguish the selected reference genomes. Tersect Browser+ also offers barcode generation functionality for various cultivars and specific species in the dataset.

\begin{figure} [http]
  \centering
  \includegraphics[width=0.5\textwidth]{path/to_image.png}
  \caption{Tersect Browser+ overview. The browser provides a single way to analyse variant data without leaving the page.[More description]}
  \label{fig:your_label}
\end{figure}

\subsubsection{Heatmap and Phylogenetic Tree}

Visualising introgression distances for each accession against the reference genome provides cultivar biologists with insights into genetic similarity, divergence patterns, and potential introgressed regions that may influence desirable traits. Tersect Browser displayed a phylogenetic tree and corresponding heatmap for the selected reference genome, chromosome, chromosomal interval, and bin size. The heatmap represented the genetic distances between each accession and the reference genome (Kurowski, 2023).
Tersect Browser+ extends and enhances these capabilities

\subsubsection{Gene Browser and Model}

Making it easier for users to explore gene models and genetic variants for a specific accession, this feature provides two ways to view the gene model track and variant track for the selected accession.  The user can either select the ‘View Gene-Model’ button to load these tracks above the accession heatmap [Ref Technical Document Figure] or right-click on a heatmap region of interest to display the accession name and open the JBrowse interface [Add reference figure from the Technical document].  This feature can also be used to [Add any other information on how this can be used].
Gene model annotations are shown in the track [Add Technical Documentation figure reference]. The paths are sorted and a compressed GFF file, along with its corresponding Tabix index are specified to the local server URLs during the browser start-up.

High-impact variants, such as frameshift mutations, premature stop codons, or major missense mutations, can have profound effects on gene function. In crop breeding programs leveraging introgression, these variants are particularly important for assessing the functional consequences of transferred genomic regions. By enabling users to search for high-impact variants based on gene ID and impact severity, Tersect Browser+ facilitates the identification of critical genetic changes associated with desirable or undesirable traits. Selected variants are visually mapped onto the heatmap, highlighting the corresponding accession and genomic region [Figure from Technical Document reference].

Providing a DNA barcode or fingerprint of the genomic composition of each accession enables rapid identification of regions derived from donor versus recipient genomes. This allows breeders to verify the successful integration of target traits and to track the use of parental lines, which is useful in identifying specific accessions in the laboratory or published datasets. In recent years, significant advances have been made in DNA barcoding methodologies. For example, Prieto et al. (2024)\cite{espinosa_prieto_finding_2024} demonstrated its application in DNA fingerprinting of tomato breeding lines, while Al-Shammari et al. (2021) \cite{al2021genetic} documented its use in cultivar investigations. In Tersect Browser+, users can generate barcodes of their preferred length for any region on a selected chromosome. The tool identifies key characteristics of interest within the barcode region, including SNP counts and positions, the number of repeat elements, and GC content [Ref figure in technical documentation showing output]. A default barcode length of 150 bp is provided, reflecting the attractiveness of lengths between 150–500 bp for primer production, as discussed by Letsiou et al. (2024) \cite{letsiou_dna_2024}.

\subsubsection{Datasets}

The Sol Genomics Network (SGN) tomato reference genome and associated resequenced accessions from cultivated lines and wild relatives was used. Reference genome updates (e.g., SL2.40 to SL2.50) were managed using SeqRemap, available under the MIT license at \url{https://bitbucket.org/cranfieldbix/seqremap}.

\vspace{3mm}

\noindent[Add on SoyBean]

@@

[Tom - we see that funding sections are covered in many papers submitted to this journal. Do we need to include? If so, what do we need to include in here?]

\vspace{3mm}

\noindent\textit{Conflicts of interest}:~none declared.

\section{Supplementary Data}

Technical documentation - pdf file

\noindent\url{[githubreadme - to be added]}
